What can Issues on GitHub and Google Code tell us about Bug Hazards in Android and Java?

### 2015-02-13 android-stacks.tex
all over

%\begin{tiny}
%\begin{verbatim}

%03-01 W/dalvikvm(7924): threadid=1: thread exiting with uncaught exception (group=0x40adf210)
%03-01 15:55:01.609 (7924): FATAL EXCEPTION: main
%03-01  (7924): java.lang.RuntimeException: ...
%E/AndroidRuntime( 7924):  at android....performLaunchActivity(ActivityThread.java:1967)
%03-01 15:55:01.609 (7924):  at android....handleLaunchActivity(ActivityThread.java:1992)
%03-01 15:55:01.609:  at android.app.ActivityThread.access\$600(ActivityThread.java:127)
%E/AndroidRuntime( 7924):  at android....handleMessage(ActivityThread.java:1158)
%\end{verbatim}
%\end{tiny}

As mentioned before, the tool enables a set of exception stack traces analyses by combining exception stack 
trace information with bytecode and sourcecode analysis. Each analysis is implemented as a different module 
on ExceptionMiner. Next, we present core analyses implemente on ExceptionMiner. Such analyses 
provide additional information to be used on further analyses described in Section~\ref{sec:result}:

\emph{\textbf{Exception Type Analysis.}} To support a deeper investigation of the 
stack traces every exception defined on a stack trace need to be classified according to its type
(e.g. Error, Exception or RuntimeException). The module responsible for this analysis 
uses Design Wizard~\cite{Brunet09} framework to inspect the bytecode of all exceptions reported on issues,
and a specific implementation to analise the source code of Java Exception classes.
It walks up the type hierarchy of a Java exception until it reaches a base exception type.
This module analyzed all exceptions defined on Android platform (API level 19) - 
which includes all basic Java exceptions that can be thrown during the app execution,
and exceptions thrown by Android core libraries. Moreover it also analized the 
libraries and applications which defined exceptions reported on stack traces
(the tool only analyzed the last version available).

\emph{\textbf{Exception Signaler Analysis.}}
This module is responsible for classifying each signaler according 
to its origin (i.e. Android Application Framework, Android Libcore, Application, Library). 
Table~\ref{tab:signalers} presents the heuristic adopted in this classification.
To enable such classification, we needed to feed this module the information
comprising all Java packages that compose: the Android Plaform;
 the Android Libcore; and each analyzed Application. 
To discover the packages composing the first two origins
we could refer to the Android Specification.
To discover the packages composing each application, this module 
extracted the manifest files of each Android app
 - which defines the main packages that compose the applications.
 Whenever this file was not available, the tool recursively analyzed the 
structure of source-code directories composing the app
- filtering out the cases on which the application 
also included the source-code of reused libraries.
Then based on this information and using pattern matching 
between the signaler name and the packages this module identifies 
the origin of the exception signalers. The exceptions were considered
 from libraries if their packages were not defined 
on Android platform, or on core libraries or on applications.
Table~\ref{tab:signalers} summarizes the signaler classification.

@@

\subsection{Manual Inspections}
In the context of the ExceptionMiner tool, manual inspections were used to:
(i) support on the identification of packages composing the Android platform, 
libs and apps analyzed in this study (as described previously); and (ii)  
identify the type of some exceptions reported on issues, 
that were not be automatically identified by the ExceptionMiner tool
(because they were defined on previous versions of libraries,
apps and Android Plaform). When the exception could not be  
found niether automatically, nor manually (because they were related an app or lib the previous versions
that were no more availble) the exception were classified 
in our study as "Undefined".  Only 31 exceptions 
remained undefined - such exceptions were reported on 61 exception stack traces as 
presented in Table ~\ref{tab:typeroottab}.

@@

 The NullPointerException was mainly signaled inside the application code (50\%) and the Android platform (31.5\%),
 although we could also find the NullPointerException being signaled by third-party libraries (16.3\%). 
Regarding reusable code (e.g., libraries and frameworks), there is no consensus whether it is a good or bad practice to 
re-throw a NullPointerException. Some prefer to encapsulate such an exception on
IllegalArgumentException, while others~\cite{bloch2008effective} argue that the
NullPointerException makes the cause of the problem explicit and hence 
should not be wrapped.

The high prevalence of NullPointerException is aligned with the findings of other research
works~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher}. For instance, Sunghun et
al.~\cite{kim2013predicting} showed that 38\% the bugs related to exception handling in Eclipse project
 were caused by NullPointerException; other works on robustness 
testing~\cite{maji2012empirical,csallner2004jcrasher} showed that most of the automatically 
detected bugs were due to NullPointerException and exceptions implicitly-signaled by Java 
environment due to programming mistakes or resource limitations
 (as the ones found in our study).

%Commented 2015
%\emph{\textbf{The prevalence of programming mistakes on stack traces.}} One way of getting a broader view of the prevalence~\footnote{The prevalence evaluates the proportion of a given characteristic over a group %of subjects.}  of programming mistakes on reported stack traces would be to manually inspect the code responsible for throwning every exception reported  on issues. However, inspecting the source code related to %6005 stack traces can easily become infeasible.
%Hence, we conducted an empirical investigation based on the following assumption: the name of the exception and its associated Javadoc 
%documentation hold enough information to pinpoint a ``programming mistake".To exemplify, an instance of ArrayOutOfBoundException
%pinpoints to programming mistake which indicates that an array has been accessed with an illegal index.

%To perform this empirical investigation, we selected a subset of all reported root exceptions. 
%This subset was composed by the top 100 root exceptions (from a list ranked by the number of ocurrences on stacks).
%This subset was representative since it contained root exceptions that were reported as the root causes of 95\% of 
%all stack traces. Therefore, based on the inspection of the full names exceptions and the associated 
%Javadocs we could identify from this subset, which exceptions were related to programming mistakes.
% It is worth mentioning that  all exceptions identified as programming mistakes were exceptions defined on java.lang and java.util packages. 
%After doing such analysis, we could calculate the prevalence of such exceptions on issues. In other words, we could calculate the proportion of issues mentioning the exceptions identified in this analysis as related to %``programming mistakes".

@@

\emph{\textbf{Identifying the Concerns Related to Root Exceptions.}} To get a broader view of the root exceptions of stack traces,
we performed an empirical evaluation whose goal was to identify the concerns related to the most reported root exceptions.
Since it is infeasible to inspect the code responsible for throwning every exception reported in this study.
This identification performed in this study was based on meaning that the name of the exception holds,
its description (i.e., Javadoc documentaion), and the Java specification. To exemplify, an instance of ArrayOutOfBoundException
usually refers to a programming mistake. Moreover, the Java specification list all exceptions related to backward compatibility
issues ~\cite{javaback} (e.g., InstantiationError, VerifyError, IllegalAccessError). Some concerns are known as sources
of faults in mobile development are: concurrency ~\cite{ama2012} backward compatibility
~\cite{McDon13}, security ~\cite{enck2011study,was2010} and resource management (IO, Memory, Batery) ~\cite{Zhang12}.
However, there are exceptions that cannot be directly mapped to one of such concerns. Either because they are
related to programming mistakes - as illustrated on the previously. Or because they are too general (i.e., java.lang.Exception,
java.lang.RuntimeException, java.lang.Error). Or also because they are related to other concerns (e.g., specific to a given library).

To perform such mapping, we selected a subset of all reported root exceptions. This subset was composed by 100 root exceptions
which were reported on 95\% of all stack traces analyzed in this study. Hence, based on the inspection of the source code and
Javadoc related to each exception, and Java specification, we identified the concern
releated to each root exception. Than we could calculate the prevalence of a concern on the reported stack traces using the following metric.
Given that: Stacks is the complete set of stacks found in this study, Lc is the list of root causes related to a given concern (as identified on the study),
and Stacks[Lc] the list of issues that mentions the concern. The prevalence of a concern could be given by the formula: Stacks[Lc]/Stacks.
Table~\ref{tab:tophundrend} illustrates the result of this empirical evaluation.

\begin{table}
\centering
\begin{tabular}{lrr}
\hline
\bfseries{Category} & \bfseries{\% Occurrences on stacks} \\
\hline
Programming logic (java.lang and util) & 52,0\%\\
Resources (IO, Memory, Batery) & 23,9\% \\
Security & 4,1\%\\
Concurrency & 2,9\% \\
Backward compatibility & 5,5\% \\
Specific Exceptions & 4,9\%\\
General (Error, Exception, Runtime) & 6,7\%\\
\end{tabular}
\caption{Identifying the concerns related to root exceptions}
\label{tab:tophundrend}
\end{table}
To ensure the quality of the process, three independent coders classified a randomly selected
sample of 25 exception types (from the total 100) using the same list of concerns;
the interater agreement was high (96\%). We could also observe that
We can therefore assume that the categories are descriptive
enough for covering the range of exception types analyzed.
These study showed that the programming mistakes associated to the most reported root causes.
However, they may be interactly related to one of the other concerns. For instance, a concurrency problem may
lead to a NullPointerException.
\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
A multitude of programming mistakes were reported as the main causes of exception stack traces.
51,96\% of all projects reported at least one exception stack trace on which the NullPoiterException
was the root cause.
}
\end{minipage}}

\subsection{What can the Types of Root Exceptions tell us?}
%\subsection{Exploring the Information Embedded on Exception Types}
%\subsection{The Exception Types and What They can Tell us? }

@@

We can see, from Table ~\ref{tab:typeroottab}, that contrasting with the other origins, most of the 
 exceptions signaled on Android Libcore (i.e., the set of libraries reused by the Android) were 
checked exceptions. Such set comprises: org.apache.harmony, org.w3c.dom, sun.misc, 
org.apache.http, org.json, org.xml, javax. Signaling checked exceptions is considered a 
good practice (see best practice IV in Section~\ref{sec:best}) because by using 
checked exceptions a library can define a precise 
exception interface ~\cite{miller1997issues} to its clients.
 Since such libraries are heavily used  in several projects, such finding can be related to the libraries maturity.

@@

We can also observe that most of the reported exceptions were runtime 
(64.85\%); and most of them were signaled on methods defined either on the Application (47.3\%)
or on the Android platform (34.3\%). We could also find runtime exceptions thrown by library code (17.7\%).
 Taking into account the results of empirical study on the prevalence of ``programming mistakes", 
we conclude that a considerable amount 
of such exceptions, are related to the programming mistakes
as pointed out previously. However, within this set of runtime exceptions there are also the ones that do not correspond to programming mistakes,
and according to Java specification should be documented as part of the exception interface of libraries/framework reusable
methods.

\emph{\textbf{Inspecting Exception Interfaces.}} To investigate the obedience to such practice, we firstly filtered out all the exceptions implicitly
 signaled by  runtime environment (due to programming mistakes) since these exceptions 
should not be documented on the method signature.  Then we inspected the code each method 
(defined either on Android  Aplication Framework or on libraries) 
explicitly signaling a runtime exception not-related to programming mistakes.
Table~\ref{tab:runtimeinterface} presents the results of this inspection. 
We could find 79 methods (both from libraries and Android platform) that  explicitly threw a runtime exception 
without listing such exception on the exception interface (i.e., using 
throws clause on method signature). From this set only one method (defined on a library)
included an @throws tag on Javadoc - reporting that the given runtime exception
could be thrown in some contidions. These methods were responsible for 
267 exception stack traces mined in this study.

This result is aligned with the results of other study conducted by  
Sacramento et al ~\cite{sacramento2006unchecked} which observed that the
runtime exceptions in .Net programs are most often not documented.
These ``undocumented runtime exceptions" represent a threat to system robustness, specially
when such exceptions are thrown by a third party code invoked inside the application 
(e.g. libraries, or framework utility code). In such cases the developer usually does not have access to 
the source code. Hence in the absence of the exception documentation it is very difficult or even impossible
 for the client to design the application to deal with ``unforeseen" runtime exceptions. As a consequence, the
 undocumented runtime exception may remain uncaught and lead to system crashes.

Moreover, these stack traces could also pointed out another best practice violation: the methods from Android platform
 presented in Table~\ref{tab:runtimeinterface} signaled direct instances of RuntimeException. 
Using such general exception, is considered a bad practice according to Java specification and compiled guidelines.

@@

 %\begin{tabular}{|p{7.9cm} p{0.1cm}|}
\begin{tabular}{lrr}
    \hline
 \bfseries{Origin}   & \bfseries{methods} &  \bfseries{stacks}  \\ 
\hline					
Libraries	& 	44	& 	205	\\					
\emph{android packages}   &  &   \\ 
android.app	& 	5	& 	16	\\
android.content	& 	2	& 	2	\\
android.database	& 	1	& 	1	\\
android.graphics	& 	5	& 	11	\\
android.hardware	& 	4	& 	5	\\
android.media	& 	3	& 	3	\\
android.opengl	& 	1	& 	8	\\
android.os	& 	11	& 	12	\\
android.text	& 	1	& 	1	\\
android.video	& 	1	& 	1	\\
android.providers	& 	1	& 	2	\\
\hline					
\bfseries{All android} 	&	35	&	62	\\
\hline					
\bfseries{All}	&	79	&	267	\\
\hline					
  \end{tabular}
\caption{Absence of exception interfaces on methods.}
\label{tab:runtimeinterface}
\end{table}

@@

\emph{\textbf{Runtime Exception wrapping Error.}} From Table~\ref{tab:wrappingandroid}, we could observe that most of such wrappings
were performed by Android platform (50.7\%). Inspecting each of them revealed that the android.os.AsyncTask, was responsible for 44\% 
of these wrappings. This class is responsible for dealing with asynchronous tasks in Android.
The code snippet below shows a general catch clause that converts any instance of Throwable (signaled during the execution
of an asynchronous task) into an instance of RuntimeException and re-throw it.  
Table~\ref{tab:exampeswrap} exemplifies some of the wrappings performed by this tasks, such as 
RuntimeExceptions wrapping istances of OutOfMemoryError and StackOverflowError. 
Such wraps mask an unrecoverable error wrapping on a general runtime exception.

@@

\emph{\textbf{Error wrapping Runtime.}} The source code inspection related that
these wrappings were performed by: native methods defined by 
java.lang.Class (e.g., getDeclaredFields, getDeclaredConstructors), 
methods from javax.crypto package. Table~\ref{tab:exampeswrap}  illustrates
examples of such wrappings. In these examples, wapping a runtime exception in an error
reflected a design decision.

\emph{\textbf{Error wrapping Checked Exception.}} 
The source code inspection related that to such wrappings were 
performed by Java static initializers. If any exception is thrown in the context of a static initializer 
(i.e., static block)  it is converted into an ExceptionInitializerError 
on the point where the class is first used.  Table~\ref{tab:exampeswrap} illustrates
examples of such wrappings. We could observe that these examples also reflected
a design decision.

@@

\subsection{On the null pointer problem}
The null references was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his ``one-billion-dollar mistake"~\cite{hoare2}:
\emph{``I call it my billion-dollar mistake. It was the invention of the null reference in 1965 [...] This has led to innumerable errors, vulnerabilities, and system 
crashes, which have probably caused a billion dollars of pain and damage in the last forty years."}

In this sudy, the null references were, infact, responsible for several reported application errors - providing more evidences to Hoare's speech.
This observation emphasizes the importance of solutions to avoid NullPointerExceptions, such as:
(i) lightweight intra-method null pointer analysis supported by Java8 @Nullable annotations~\footnote{They are used by tools e.g., Eclipse, IntelliJ, Android Studio 0.5.5 (release Apr. 2014) to detect potential 
null pointer dereferences at compile time.};
(ii) inter-method null pointer analysis tools such as the one proposed by Nanda and Sinha ~\cite{nanda2009accurate};
or (iii) language designs which avoid null pointers, such 
as Monads ~\cite{Walde95} (i.e., used in functional languages for values that may not be available 
or computations that may fail) could improve the robustness of Java programs.

@@

\textit{Analysis and Use of Stack Trace Information.} Several research works have
investigated the use of stack trace information to support bug classification
and clustering~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},
fault-proneness prediction models~\cite{kim2013predicting} and even automated
bug fixing tools~\cite{sinha2009fault}. Kim et al.~\cite{kim2011crash} use an
aggregated form of multiple stack traces available in crash reports to detect
duplicate crash reports and to predict if a given crash will be fixed. Dhaliwal
et al.~\cite{dhaliwal2011classifying} proposed a crash grouping approach that
can reduce bug fixing time in approximately 5\%. Wang et
al.~\cite{wang2013improving} propose an approach to identify correlated crash
types and describe a fault localization method to locate and rank files related
to the bug described on a stack trace. Schroter et al.~\cite{schroter2010stack}
conducted an empirical study on the usefulness of stack traces for bug fixing
and showed that developers fixed the bugs faster when failing stack traces were
included on bug issues.  In a similar study, Bettenburg et
al.~\cite{bettenburg2008makes} identify stack traces as the second most stack
trace feature for developers.  Sinha et al.~\cite{sinha2009fault} proposed an
approach that uses stack traces to guide a dataflow analysis for locating and
repairing faults that are caused by the implicitly signaled exceptions. Kim
at al.~\cite{kim2013predicting} proposed an approach to predict the
crash-proneness of methods based information extracted from stack traces and
methods' bytecode operations.  They observed that most of the stack traces were
related to NullPointerException and other implicitly thrown exceptions had
the higher prevalence on the analyzed set of stacks. Kechagia and Spinellis ~\cite{kechagia2014}
examined the stack traces embedded on crash reports sent by 1,800 Android apps 
to a crash report management service (i.e., BugSense). They found that 19% of such stack traces
were caused by unchecked and undocumented exceptions thrown by methods defined on 
Android API (level 15). Our work differs from Kechagia and Spinellis since it is not based on
stack traces mined from issues reported by open source developers on GitHub and Googlecode.  
Moreover, our work identified the type of each exception mined from issues
(classifying them as Error, Runtime or Checked) based on the source code
analysis of the exception hierarchy; we mapped the exception signalers (i.e., were the
exceptions signaled from libraries, the platform or the application itself); we investigated
the adoption of best practices based on the analysis of stack trace information.
Such analysis revealed intrigging cross-type exception wrappings which may bring the program 
to an unpredictable state (i.e.,  trying to handle an instance of OutOfMemoryError “hidden” in a checked exception).

>>>>

\emph{\textbf{Exception Type Analysis.}} To support a deeper investigation of the 
stack traces every exception defined on a stack trace need to be classified according to its type
(e.g. Error, Exception or RuntimeException). The module responsible for this analysis 
uses Design Wizard~\cite{Brunet09} framework to inspect the bytecode of all exceptions reported on issues,
and a specific implementation to analyze the source code of Java Exception classes (when the
bytecode was not available). It walks up the type hierarchy of a Java exception until it reaches a base exception type.
This module analyzed all exceptions defined on Android platform (API level 19) - 
which includes all basic Java exceptions that can be thrown during the app execution,
and exceptions thrown by Android core libraries. Moreover it also analyzed the 
libraries and applications which defined exceptions reported on stack traces
(the tool only analyzed the last version available).

\emph{\textbf{Exception Signaler Analysis.}}
This module is responsible for classifying each signaler according 
to its origin (i.e. Android Application Framework, Android Libcore, Application, Library). 
Table~\ref{tab:signalers} presents the heuristic adopted in this classification.
To enable such classification, we needed to feed this module the information
comprising all Java packages that compose: the Android Plaform;
 the Android Libcore; and each analyzed Application. 
To discover the packages composing the first two origins
we could refer to the Android Specification.
To discover the packages composing each application, this module 
extracted the manifest files of each Android app
 - which defines the main packages that compose the applications.
 Whenever this file was not available, the tool recursively analyzed the 
structure of source-code directories composing the app
- filtering out the cases on which the application 
also included the source code of reused libraries.
Then based on this information and using pattern matching 
between the signaler name and the packages this module identifies 
the origin of the exception signalers. The exceptions were considered
 from libraries if their packages were not defined 
on Android platform, or on core libraries or on applications.
Table~\ref{tab:signalers} summarizes the signaler classification.

@@

\subsection{Manual Inspections}
In the context of the ExceptionMiner tool, manual inspections were used to:
(i) support on the identification of packages composing the Android platform, 
libs and apps analyzed in this study (as described previously); and (ii)  
identify the type of some exceptions reported on issues, 
that were not be automatically identified by the ExceptionMiner tool
(because they were defined on previous versions of libraries,
apps and Android Plaform). When the exception could not be  
found niether automatically, nor manually (because they were defined on a previous version
of the app or lib) the exception were classified 
in our study as "Undefined".  Only 31 exceptions 
remained undefined - such exceptions were reported on 61 exception stack traces as 
presented in Table ~\ref{tab:typeroottab}.

@@

 The NullPointerException was mainly signaled inside the application code (50\%) and the Android platform (31.5\%),
 although we could also find the NullPointerException being signaled by third-party libraries (16.3\%). 
Regarding reusable code (e.g., libraries and frameworks), there is no consensus whether it is a good or a bad practice to 
re-throw a NullPointerException. Some prefer to encapsulate such an exception on
an instance of IllegalArgumentException, while others~\cite{bloch2008effective} argue that the
NullPointerException makes the cause of the problem explicit and hence 
should not be wrapped.

The high prevalence of NullPointerException is aligned with the findings of other research
works~\cite{kim2013predicting,fraser20131600,csallner2004jcrasher,kechagia2014}. For instance, Sunghun et
al.~\cite{kim2013predicting} showed that 38\% the bugs related to exception handling in Eclipse project
 were caused by NullPointerException; Kechagia and Diomidis showed that NullPointerException was the  
most reported exception on the crash reports sent to BugSense (a bug report 
management service for Android applications); 
other works on robustness testing~\cite{maji2012empirical,csallner2004jcrasher} showed that most of the automatically 
detected bugs were due to NullPointerException and exceptions implicitly-signaled by Java 
environment due to programming mistakes or resource limitations
 (as the ones found in our study).

%Commented 2015
%\emph{\textbf{The prevalence of programming mistakes on stack traces.}} One way of getting a broader view of the prevalence~\footnote{The prevalence %evaluates the proportion of a given characteristic over a group %of subjects.}  of programming mistakes on reported stack traces would be to manually %inspect the code responsible for throwning every exception reported  on issues. However, inspecting the source code related to %6005 stack traces can %easily become infeasible.
%Hence, we conducted an empirical investigation based on the following assumption: the name of the exception and its associated Javadoc 
%documentation hold enough information to pinpoint a ``programming mistake".To exemplify, an instance of ArrayOutOfBoundException
%pinpoints to programming mistake which indicates that an array has been accessed with an illegal index.

%To perform this empirical investigation, we selected a subset of all reported root exceptions. 
%This subset was composed by the top 100 root exceptions (from a list ranked by the number of ocurrences on stacks).
%This subset was representative since it contained root exceptions that were reported as the root causes of 95\% of 
%all stack traces. Therefore, based on the inspection of the full names exceptions and the associated 
%Javadocs we could identify from this subset, which exceptions were related to programming mistakes.
% It is worth mentioning that  all exceptions identified as programming mistakes were exceptions defined on java.lang and java.util packages. 
%After doing such analysis, we could calculate the prevalence of such exceptions on issues. In other words, we could calculate the proportion of issues %mentioning the exceptions identified in this analysis as related to %``programming mistakes".

@@

\emph{\textbf{Identifying the Concerns Related to Root Exceptions.}} To get a broader view of the root exceptions of stack traces,
we performed an empirical evaluation whose goal was to identify the concerns related to the most reported root exceptions.
Besides the exceptions related to programming mistakes mentined before, we also looked for exceptions related to some concerns that are known as sources of faults in mobile development: concurrency ~\cite{ama2012} backward compatibility ~\cite{McDon13}, security ~\cite{enck2011study,was2010} and resource management (IO, Memory, Batery) ~\cite{Zhang12}.

Since it is infeasible to inspect the code responsible for throwning every exception reported in this study,
the concern identification of each exception was based on meaning of the exception type defined on 
its Javadoc documentaion and on the Java specification. To exemplify: (i) an instance of ArrayOutOfBoundException  
refers to a programming mistake according to its Javadoc; and (ii) the Java specification list all exceptions related to backward compatibility
%issues ~\cite{javaback} (e.g., InstantiationError, VerifyError, IllegalAccessError).

To perform such mapping, we selected a subset of all reported root exceptions. This subset was composed by 100 root exceptions
which were reported on 95\% of all stack traces analyzed in this study. Hence, based on the inspection of the 
Javadoc related to each exception, and Java specification, we identified the concern releated to each root exception. 
Table~\ref{tab:tophundrend} illustrates the result of this empirical evaluation. This table also illustrates the exceptions 
that could not be directly mapped to one of the aforementioned concerns, either because they were too general (i.e., java.lang.Exception,
java.lang.RuntimeException, java.lang.Error) or because they were related to other concerns (e.g., specific to an application or given library).

\begin{table}
\centering
\begin{tabular}{lrr}
\hline
\bfseries{Category} & \bfseries{\% Occurrences on stacks} \\
\hline
Programming logic (java.lang and util) & 52,0\%\\
Resources (IO, Memory, Batery) & 23,9\% \\
Security & 4,1\%\\
Concurrency & 2,9\% \\
Backward compatibility & 5,5\% \\
Specific Exceptions & 4,9\%\\
General (Error, Exception, Runtime) & 6,7\%\\
\end{tabular}
\caption{Identifying the concerns related to root exceptions}
\label{tab:tophundrend}
\end{table}

To ensure the quality of the process, three independent coders classified a randomly selected
sample of 25 exception types (from the total 100) using the same list of concerns;
the interater agreement was high (96\%). We could also observe that
We can therefore assume that the categories are descriptive
enough for covering the range of exception types analyzed.

%These study showed that the programming mistakes associated to the most reported root causes.
%However, they may be interactly related to one of the other concerns. For instance, a concurrency problem may
%lead to a NullPointerException.

%For instance, the Java specification list all exceptions related to backward compatibility
%issues ~\cite{javaback} (e.g., InstantiationError, VerifyError, IllegalAccessError). 
%were also investigated in this study

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
A multitude of programming mistakes were reported as the main causes of exception stack traces.
52\% of all projects reported at least one exception stack trace on which the NullPoiterException
was the root cause.
}
\end{minipage}}

\subsection{What can the Exception Types tell us?}
%\subsection{Exploring the Information Embedded on Exception Types}
%\subsection{The Exception Types and What They can Tell us? }

@@

We can observe that most of the reported exceptions were runtime 
(64.85\%); and most of them were signaled on methods defined either on the Application (47.3\%)
or on the Android platform (34.3\%). We could also find runtime exceptions thrown by library code (17.7\%).

We can also see, from Table ~\ref{tab:typeroottab}, that contrasting with the other origins, most of the 
 exceptions signaled on Android Libcore (i.e., the set of libraries reused by the Android) were 
checked exceptions. Such set comprises: org.apache.harmony, org.w3c.dom, sun.misc, 
org.apache.http, org.json, org.xml, javax. Signaling checked exceptions is considered a 
good practice (see best practice IV in Section~\ref{sec:best}) because by using 
checked exceptions a library can define a precise 
exception interface ~\cite{miller1997issues} to its clients.
 Since such libraries are heavily used  in several projects, such finding can be related to the libraries maturity.

@@

\emph{\textbf{Inspecting Exception Interfaces.}}  According to the best practices mentioned before, runtime exceptions 
should be documented as part of the exception interface of libraries/framework reusable
methods. To investigate the obedience to such practice, we firstly filtered out all the exceptions implicitly
 signaled by  runtime environment (due to programming mistakes) - since these exceptions 
should not be documented on the method signature.  Then we inspected the code each method 
(defined either on Android  Aplication Framework or on third-party libraries) 
explicitly signaling a runtime exception not-related to programming mistakes.
Table~\ref{tab:runtimeinterface} presents the results of this inspection. 
We could find 79 methods (both from libraries and Android platform) that  explicitly threw a runtime exception 
without listing such exception on the exception interface (i.e., using 
throws clause on method signature). From this set only one method (defined on a library)
included an @throws tag on Javadoc - reporting that the given runtime exception
could be thrown in some contidions. These methods were responsible for 
267 exception stack traces mined in this study.

This result is aligned with the results of other study conducted by  
Sacramento et al ~\cite{sacramento2006unchecked} - which observed that the
runtime exceptions in .Net programs are most often not documented; and 
Kechagia and Spinellis ~\cite{kechagia2014} which discovered a set of methods 
on the Android API which do not documented runtime exceptions. One limitation 
of the work of Kechagia and Spinellis is that it did filter out exceptions that should 
although unchecked not be documented -  because they are implicitly signaled by the 
JVM due to resource restrictions or violations on Java semantic constraints.

The ``undocumented runtime exceptions" that are explicitly siglened represent a threat to system robustness, specially
when such exceptions are thrown by a third party code invoked inside the application 
(e.g. libraries, or framework utility code). In such cases the developer usually does not have access to 
the source code. Hence in the absence of the exception documentation it is very difficult or even impossible
 for the client to design the application to deal with ``unforeseen" runtime exceptions. As a consequence, the
 undocumented runtime exception may remain uncaught and lead to system crashes.

@@

 %\begin{tabular}{|p{7.9cm} p{0.1cm}|}
\begin{tabular}{lrr}
    \hline
 \bfseries{Origin}   & \bfseries{methods} &  \bfseries{stacks}  \\ 
\hline					
Libraries	& 	44	& 	205	\\
Android  	&	35	&	62	\\					
%\emph{android packages}   &  &   \\ 
%android.app	& 	5	& 	16	\\
%android.content	& 	2	& 	2	\\
%android.database	& 	1	& 	1	\\
%android.graphics	& 	5	& 	11	\\
%android.hardware	& 	4	& 	5	\\
%android.media	& 	3	& 	3	\\
%android.opengl	& 	1	& 	8	\\
%android.os	& 	11	& 	12	\\
%android.text	& 	1	& 	1	\\
%android.video	& 	1	& 	1	\\
%android.providers	& 	1	& 	2	\\
%\hline					
%\bfseries{All android} 	&	35	&	62	\\
\hline					
\bfseries{All}	&	79	&	267	\\
\hline					
  \end{tabular}
\caption{Absence of exception interfaces on methods.}
\label{tab:runtimeinterface}
\end{table}

@@

\emph{\textbf{Runtime Exception wrapping Error.}} From Table~\ref{tab:wrappingandroid}, we could observe that most of such wrappings
were performed by Android platform (50.7\%). Inspecting each of them revealed that the android.os.AsyncTask, was responsible for 44\% 
of these wrappings. This class is responsible for dealing with asynchronous tasks in Android.
The code snippet below shows a general catch clause that converts any instance of Throwable (signaled during the execution
of an asynchronous task) into an instance of RuntimeException and re-throw it.  
Table~\ref{tab:exampeswrap} exemplifies some of the wrappings performed in this code snipet:  
an instance of RuntimeExceptions wrapping instances of OutOfMemoryError and StackOverflowError. 
Such wraps mask an unrecoverable error wrapping on a general runtime exception.

Moreover, such wrapping also pointed out to another best practice violation: the methods from Android platform
signaling direct instances of RuntimeException. Using such general exception, is considered a bad practice 
according to Java specification and compiled guidelines, because it loses contextual information about the exception.

@@

\emph{\textbf{Error wrapping Runtime and Checked Exceptions}} Table~\ref{tab:exampeswrap}  illustrates
examples of instances of Error wrapping instances of RuntimeException. A
lthough such wapping mixes different exception types,
since there is no obligation associated to handling runtime exceptions, such wrapping 
does not violated the aforementioned best practices. On the other hand, the inspection 
also revealed instances of Error wrapping checked exceptions. Such wrappings were 
mostly performed by Java static initializers. If any exception is thrown in the context of a static initializer 
(i.e., static block)  it is converted into an ExceptionInitializerError 
on the point where the class is first used.  Table~\ref{tab:exampeswrap} also illustrates
examples of such wrappings. Although such wapping may represent a design decision,
it violates the best practice related to checked exceptions and errors as it mix the intended handling 
behaviour associated to both types.

@@

\subsection{On the null pointer problem}
The null references was firstly introduced by Tony Hoare in ALGOL W, which after some years he called his ``one-billion-dollar mistake"~\cite{hoare2}.

%:
%\emph{``I call it my billion-dollar mistake. It was the invention of the null reference in 1965 [...] 
%This has led to innumerable errors, vulnerabilities, and system 
%crashes, which have probably caused a billion dollars of pain and damage in the last forty years."}

In this sudy, the null references were, infact, responsible for several reported issues - providing more evidences to Hoare's speech.
This observation emphasizes the importance of solutions to avoid NullPointerExceptions, such as:
(i) lightweight intra-method null pointer analysis supported by Java8 @Nullable annotations~\footnote{They are used by tools e.g., Eclipse, IntelliJ, Android Studio 0.5.5 (release Apr. 2014) to detect potential 
null pointer dereferences at compile time.};
(ii) inter-method null pointer analysis tools such as the one proposed by Nanda and Sinha ~\cite{nanda2009accurate};
or (iii) language designs which avoid null pointers, such 
as Monads ~\cite{Walde95} (i.e., used in functional languages for values that may not be available 
or computations that may fail) could improve the robustness of Java programs.

@@

\textit{Analysis and Use of Stack Trace Information.} Several research works have
investigated the use of stack trace information to support bug classification
and clustering~\cite{wang2013improving, kim2011crash, dhaliwal2011classifying},
fault-proneness prediction models~\cite{kim2013predicting} and even automated
bug fixing tools~\cite{sinha2009fault}. Kim et al.~\cite{kim2011crash} use an
aggregated form of multiple stack traces available in crash reports to detect
duplicate crash reports and to predict if a given crash will be fixed. Dhaliwal
et al.~\cite{dhaliwal2011classifying} proposed a crash grouping approach that
can reduce bug fixing time in approximately 5\%. Wang et
al.~\cite{wang2013improving} propose an approach to identify correlated crash
types and describe a fault localization method to locate and rank files related
to the bug described on a stack trace. Schroter et al.~\cite{schroter2010stack}
conducted an empirical study on the usefulness of stack traces for bug fixing
and showed that developers fixed the bugs faster when failing stack traces were
included on bug issues.  In a similar study, Bettenburg et
al.~\cite{bettenburg2008makes} identify stack traces as the second most stack
trace feature for developers.  Sinha et al.~\cite{sinha2009fault} proposed an
approach that uses stack traces to guide a dataflow analysis for locating and
repairing faults that are caused by the implicitly signaled exceptions. Kim
at al.~\cite{kim2013predicting} proposed an approach to predict the
crash-proneness of methods based information extracted from stack traces and
methods' bytecode operations.  They observed that most of the stack traces were
related to NullPointerException and other implicitly thrown exceptions had
the higher prevalence on the analyzed set of stacks. Kechagia and Spinellis ~\cite{kechagia2014}
examined the stack traces embedded on crash reports sent by 1,800 Android apps 
to a crash report management service (i.e., BugSense). They found that 19% of such stack traces
were caused by unchecked and undocumented exceptions thrown by methods defined on 
Android API (level 15). Our work differs from Kechagia and Spinellis since it is not based on
stack traces mined from issues reported by open source developers on GitHub and Googlecode.  
Moreover, our study mapped the origin of each exception 
(i.e., libraries, the Android platform or the application itself) and investigated
the adoption of best practices based on the analysis of stack trace information.
Our work also identified the type of each exception mined from issues
(classifying them as Error, Runtime or Checked) based on the source code
analysis of the exception hierarchy and analized the exception wrappings that can 
happen during the exception propagation.  Such analysis revealed intrigging 
bug hazards such as the cross-type exception wrappings not discussed in previous works.
